A Multi-telescope Campaign on \frb: Implications for the FRB Population

### 2017-05-21 repeater-vladata.tex
through all comments but jason

@@

Many new models for the origin of FRBs have been developed in response to the localization of \frb\ \citep{2017arXiv170104815K, 2017arXiv170102370M, 2017arXiv170104094Z, 2017arXiv170102492D, 2017arXiv170208644B, 2017arXiv170300393T}. The association of \frb\ with a compact, persistent radio source is consistent with bursts coming from a young magnetar that powers a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb\ is found in a low-metallicity dwarf galaxy \citep{OPT}, which was not predicted by origin models that scale with stellar mass or star formation \citep{2017arXiv170400022N}. These galaxies are the preferred environment for long GRBs and Type-I superluminous supernovae (LGRB and SLSN-I, respectively), which have been suggested are signatures of the magnetar birth \citep{2008AJ....135.1136M, 2014ApJ...787..138L}.

@@

In Section \ref{sec:obs}, we describe the multi-telescope observing campaign and a refined analysis of the nine VLA bursts. Section \ref{sec:res} presents the first spectrum of an FRB simultaneously detected at multiple telescopes and confirm that burst spectra cannot be modeled with a single spectral index. We then model the dynamic spectra to characterize the burst spectra, energies, and dispersion properties. Section \ref{sec:disc} discusses the properties of \frb\ bursts and its impact on inferences about the birth rate of FRB sources and strategies to find and/or localize new FRBs.

@@

All VLA fast-sampled data were observed with 5~ms sampling, 256 channels, and dual-circular polarization \citep{2015ApJ...807...16L}. The total bandwidths at L (1.4~GHz), S (3~GHz), and C (6~GHz) bands were 256, 1024, and 2048~MHz, respectively, corresponding to channel bandwidths of  1, 4, and 8~MHz. The dispersion smearing across a channel is $\sim1.7$, 0.7, and 0.2~ms in the same respective bands. The 3~GHz data were recorded data in eight spectral windows with 32 channels each and had sensitivity of 5~mJy in 5~ms ($1\sigma$).

@@

Arecibo observed with the L-wide receiver using the PUPPI pulsar backend. The observational frequency range was 1.15 to 1.73~GHz and frequency resolution was 1.5625~MHz, which has a typical sensitivity of 2~mJy in 2~ms ($1\sigma$). We recorded total full Stokes polarization intensity spectra with a time resolution of 10.24~$\mu$s. Each frequency channel was coherently dedispersed to 557~pc cm$^{-3}$, significantly reducing intra-channel dispersion smearing. The full width at half maximum (FWHM) beam size at band center is 3.3\arcmin.

@@

Beginning in April 2016, VLA observations of \frb\ were simultaneously observed with LWA1 when possible.  The LWA1 observations were automatically scheduled through the Heuristic Automation for LWA1 (HAL) system.  Briefly, the HAL system receives messages via the internet that signal the start of a VLA observation. If the source is visible and there are no high priority LWA1 observations scheduled, the HAL system automatically generates an observing schedule, re-configures the telescope, and alerts the observatory staff of the change. The time delay from HAL receiving a message to starting an observation is typically one to two minutes.

@@

No dispersed pulses were found with significance greater than $10\sigma$\, equivalent to a flux density limit of $\sim60$\ Jy for a width of 160~ms. The system equivalent flux density does not change much between the two windows. The effective sensitivity to a typical 2~ms pulse width is 4.8~kJy.

@@

Figure \ref{fig:sgram} shows a dynamic spectrum for the first FRB burst to be detected simultaneously at two telescopes. This burst, on MJD 57648, was detected both by the VLA and Arecibo while upper limits were measured by simultaneous observations at the other three telescopes. In total, four of the VLA bursts had simultaneous observing coverage with either Arecibo, Effelsberg, LWA1, or AMI-LA. Two of the VLA bursts had simultaneous observing coverage by all four observatories, including the burst detected by Arecibo (Figure \ref{fig:sched}).

@@

While most bursts are modeled with well-defined parameter probability distributions, one burst is an outlier. Figure \ref{fig:corner} shows the scatter plot matrix \citep{corner} for the MCMC run on burst 57646. Two clusters of samples are identified for this burst: one narrow, low-DM and one wide, high-DM. Assuming that the model is appropriate, this implies that the burst can be decomposed into two subbursts.

@@

The spectral fits in Figure \ref{fig:spec} have residuals that include single samples that deviate by many standard deviations, particularly in the bursts on MJDs 57623,  57633.6, and 57643. These do not appear to be caused by terrestrial interference. They are most likely spectral variations from diffractive interstellar scintillations that, as noted previously, have characteristic frequency widths slightly larger than the 4~MHz channel bandwidth.  Scintillation variations are multiplicative and have a one-sided exponential probability distributino function that can yield occasional large-amplitude ``scintles'' that are several times the mean scintle amplitude of unity.

@@

The burst-specific DM structure could be intrinsic to the emission mechanism or induced during propagation. Radio waves can be modified in a variety of ways (e.g., scintillation, scattering) and the duration of the emission in the source frame is not known \citep{2016arXiv160505890C}. \citet{2017arXiv170306580C} describe a model where plasma lensing near the source of an FRB can magnify radio emission by orders of magnitude. Plasma lensing can also produce multiple burst images separated by time scales from microseconds to tens of milliseconds or longer. These burst images would also have slightly different DMs. Lensing amplifications have a complicated frequency dependence that includes spikes, plateaus, and troughs. The simultaneous detection of burst MJD 57648 with both Arecibo (L band) and the VLA (S band up to 3.2~GHz) suggests requires a focal frequency $\gtrsim 3.2$ GHz and sufficient amplification in both bands.

@@

The chance of detecting a burst from \frb\ with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects in the host galaxy or intergalactic medium \citep{2017arXiv170306580C}. However, the temporal distribution of VLA detections alone shows that the \frb\ burst rate is significantly correlated on short timescales. That is, detecting a burst predicts that you are more likely to find another burst soon thereafter and not detecting a burst predicts you will likely not detect another burst soon thereafter.

@@

Despite its many assumptions, this volumetric rate is more appropriate for comparison to rates used for other classes of transient. The LGRB and SLSN-I rates are $10^{-7}$\ and $10^{-8}$\ Mpc$^{-3}$\ yr$^{-1}$, respectively \citep{2007ApJ...657L..73G,2012Sci...337..927G}. Comparing the FRB rate to these rates implies that FRB-emitting sources can be generated in LGRBs and SLSN-I if FRBs repeat $5\times10^2$\ and $5\times10^3$\ times\footnote{\citet{2017arXiv170400022N} provide a similar estimate of FRB birth rate and reach a similar conclusion.}, respectively (assuming burst energies of $\sim10^{40}$\ erg). \citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered its magnetic field as $\approx300$. Assuming this model for FRBs, LGRBs, and SLSN-I, the rate comparison suggest that the LGRB rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate.

>>>>

\author{K.~Gourdji}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

@@

Many new models for the origin of FRBs have been developed in response to the localization of \frb\ \citep{2017arXiv170104815K, 2017arXiv170102370M, 2017arXiv170104094Z, 2017arXiv170102492D, 2017arXiv170208644B, 2017arXiv170300393T}. The association of \frb\ with a compact, persistent radio source is consistent with bursts coming from a young magnetar that powers a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb\ is found in a low-metallicity dwarf galaxy \citep{OPT}, which was not predicted by origin models that scale with stellar mass or star formation \citep{2017arXiv170400022N}. These galaxies are the preferred environment for long GRBs and Type-I superluminous supernovae (LGRB and SLSN-I, respectively), which have been suggested are signatures of magnetar birth \citep{2008AJ....135.1136M, 2014ApJ...787..138L}.

@@

In Section \ref{sec:obs}, we describe the multi-telescope observing campaign and a refined analysis of the nine VLA bursts. Section \ref{sec:res} presents the first spectrum of an FRB simultaneously detected at multiple telescopes, confirming that burst spectra cannot be modeled with a single spectral index. We then model the dynamic spectra to characterize the burst spectra, energies, and dispersion properties. Section \ref{sec:disc} discusses the properties of \frb\ bursts and their impact on inferences about the birth rate of FRB sources and strategies to find and/or localize new FRBs.

@@

All VLA fast-sampled data were observed with 5~ms sampling, 256 channels, and dual-circular polarization \citep{2015ApJ...807...16L}. The total bandwidths at L (1.4~GHz), S (3~GHz), and C (6~GHz) bands were 256, 1024, and 2048~MHz, respectively, corresponding to channel bandwidths of  1, 4, and 8~MHz. The dispersion smearing across a channel is $\sim1.7$, 0.7, and 0.2~ms in the same respective bands. The 3~GHz data were recorded in eight spectral windows with 32 channels each and had a sensitivity of 5~mJy in 5~ms ($1\sigma$).

@@

Arecibo observed with the L-wide receiver using the PUPPI pulsar backend. The observational frequency range was 1.15 to 1.73~GHz and frequency resolution was 1.5625~MHz, which has a typical sensitivity of 2~mJy in 2~ms ($1\sigma$). We recorded intnsities of the each of the two orthogonal linear polarization signals and their cross product. Full Stokes polarization intensity spectra can then be generated with a time resolution of 10.24~$\mu$s. Each frequency channel was coherently dedispersed to 557~pc cm$^{-3}$, significantly reducing intra-channel dispersion smearing. The full width at half maximum (FWHM) beam size at band center is 3.3\arcmin.

@@

Beginning in April 2016, VLA observations of \frb\ were simultaneously observed with LWA1 when possible.  The LWA1 observations were automatically scheduled through the Heuristic Automation for LWA1 (HAL) system.  Briefly, the HAL system receives messages via the internet that signals the start of a VLA observation. If the source is visible and there are no high priority LWA1 observations scheduled, the HAL system automatically generates an observing schedule, re-configures the telescope, and alerts the observatory staff of the change. The time delay from HAL receiving a message to starting an observation is typically one to two minutes.

@@

No dispersed pulses were found with significance greater than $10\sigma$, equivalent to a flux density limit of $\sim60$\ Jy for a width of 160~ms. The system equivalent flux density does not change much between the two windows. The effective sensitivity to a typical 2~ms pulse width is 4.8~kJy.

@@

Figure \ref{fig:sgram} shows a dynamic spectrum for the first FRB burst to be detected simultaneously at two telescopes. This burst, on MJD 57648, was detected both by the VLA and Arecibo while upper limits were measured by simultaneous observations at the other three telescopes. In total, four of the VLA bursts had simultaneous observing coverage with either Arecibo, Effelsberg, LWA1, or AMI-LA. Two of the VLA bursts had simultaneous observing coverage by all four observatories, including the burst detected by Arecibo.

@@

While most bursts are modeled with well-defined parameter probability distributions, one burst is an outlier. Figure \ref{fig:corner} shows the scatter plot matrix \citep{corner} for the MCMC run on burst 57646. Two clusters of samples are identified for this burst: one narrow, low-DM and one wide, high-DM. This shows that the model is not appropriate for this burst and could indicate that the burst can be decomposed into two subbursts.

@@

The spectral fits in Figure \ref{fig:spec} have residuals that include single samples that deviate by many standard deviations, particularly in the bursts on MJDs 57623,  57633.6, and 57643. These do not appear to be caused by terrestrial interference. They are most likely spectral variations from diffractive interstellar scintillations that, as noted previously, have characteristic frequency widths slightly larger than the 4~MHz channel bandwidth.  Scintillation variations are multiplicative and have a one-sided exponential probability distribution function that can yield occasional large-amplitude ``scintles'' that are several times the mean scintle amplitude of unity.

@@

The burst-specific DM structure could be intrinsic to the emission mechanism or induced during propagation. Radio waves can be modified in a variety of ways (e.g., scintillation, scattering) and the duration of the emission in the source frame is not known \citep{2016arXiv160505890C}. \citet{2017arXiv170306580C} describe a model where plasma lensing near the source of an FRB can magnify radio emission by orders of magnitude. Plasma lensing can also produce multiple burst images separated by time scales from microseconds to tens of milliseconds or longer. These burst images would also have slightly different DMs. Lensing amplifications have a complicated frequency dependence that includes spikes, plateaus, and troughs. The simultaneous detection of burst MJD 57648 with both Arecibo (L band) and the VLA (S band up to 3.2~GHz) requires a focal frequency $\gtrsim 3.2$ GHz and sufficient amplification in both bands.

@@

The chance of detecting a burst from \frb\ with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects in the host galaxy or intergalactic medium \citep{2017arXiv170306580C}. However, the temporal distribution of VLA detections alone shows that the \frb\ burst rate is significantly correlated on short timescales. That is, detecting a burst implies a higher likelihood of detecting another burst soon thereafter, and not detecting a burst also implies a higher likelihood of not detecting another burst soon thereafter.

@@

Despite its many assumptions, this volumetric rate is more appropriate for comparison to rates used for other classes of transient. The LGRB and SLSN-I rates are $10^{-7}$\ and $10^{-8}$\ Mpc$^{-3}$\ yr$^{-1}$, respectively \citep{2007ApJ...657L..73G,2012Sci...337..927G}. Comparing the FRB rate to these rates implies that FRB-emitting sources can be generated in LGRBs and SLSN-I if FRBs repeat $5\times10^2$\ and $5\times10^3$\ times\footnote{\citet{2017arXiv170400022N} provide a similar estimate of FRB birth rate and reach a similar conclusion.}, respectively (assuming burst energies of $\sim10^{40}$\ erg). \citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered by its magnetic field as $\approx300$. Assuming this model for FRBs, LGRBs, and SLSN-I, the rate comparison suggest that the LGRB rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate.
